---
title: "Comments for V. Espinoza HW-4 ES207"
author: Cininta Pertiwi
output: html_notebook
---

#### **STRINGS**

##### *String Basics number 6*

* The calling of the packages could be inside the function with `require()` instead of outside the function. That way when the function used it will call the package on its own.
* When a vector length is 2, example `c("a", "b")` this results in `a, and b`, maybe another layer in the function for if `length(x) == 2` could be specified so that the separator for a vector length of 2 is "and".

##### *Other*

* I think you did a great job Vicky! I learned from your code, especially the ones that I was struggling with to finish (Tools: Grouped & Splitting; Count most words used in sentences). I also think the codes and functions you used were simpler and there were some functions I didn't know about. An example is `intersect()` that was used in Replacing Matches numer 3. Thank you! :)

#### **AIR QUALITY DATA**

##### *Number 8*
For the code below, I think you could also use piping or make it into a function, maybe?
```{r eval = FALSE}
# code not run

x <- c("San", "Santa")
name_match <- str_c(x,collapse = "|")
name_subset <- str_subset(loc$`Site Name`,name_match)
count_names <- str_count(name_subset,name_match)
sum(count_names)
```

##### *Number 9*
The code you had (in chunk below), does this only count the missing addresses and/or zip codes? There were also values of adresses and zip codes that were not missing but did not give a complete searchable address and I was wondering if this code would also account for those.
```{r eval = FALSE}
# code not run

sum(duplicated(loc$Location)) #this is to confirm that Locations are not duplicated 
df <- tibble(
  address=loc$Address, Zip=loc$`Zip Code`
)
sum(is.na(df))
```

##### *Number 11*
You mentioned that you were not able to rerun the code. I didn't run it either but took a look at it. I made minor comments that would maybe be helpful to get the code rerunning.
```{r eval = FALSE}
# code not run

library(data.table) # This can be put into the function with 'require()'
library(tidyverse)

site.summary <- function(x,y,z){
  
  yearly <- x %>%
    group_by(site = as.factor(site))%>%
    mutate(year = format(date, "%Y")) %>%
    group_by(site,year) %>% # I'm not sure if it's possible to pipe into the newly defined c variable; Was this pipe sign supposed to be here?
    c <- c(y, z)
    
  name_match <- str_c(c,collapse = "|")
  d <- str_subset(loc$`Site Name`,name_match)
  
  San <- subset(yearly, yearly$`Site Name`== d )%>%
    group_by(site,year )%>%
    summarize(mean=mean(o3, na.rm=TRUE),median=median(o3, na.rm = TRUE),max=max(o3, na.rm=TRUE),min=min(o3, na.rm=TRUE))
  
  San # The function needs to return something; If you meant to return 'San' it probably should be 'return(San)'
}
site.mean(daily.site,"San","Santa")
```

##### *Number 12*
Similar to number 11, I didn't run the code but took a look at it. I made minor comments that would maybe be helpful to get the code rerunning.
```{r eval = FALSE}
# code not run

library(data.table) # This can be put into the function with 'require()'
library(tidyverse)

site.mean <- function(x){
  
  x <- c("Merced") # Maybe piping operations might be possible for this?
  name_match <- str_c(x,collapse = "|")
  name_subset <- str_subset(loc$`Site Name`,name_match)
  merceddata <- subset(daily.site, daily.site$`Site Name`== name_subset)
  mean=mean(x, na.rm=TRUE)
  
  # The function needs to return something. Is the function returning merceddata or mean? I don't think functions can return two things.
}
site.mean(merceddata$o3)
mean # Is this mean variable same as what was defined in the function? I'm not sure how this plays out because isn't the 'mean' variable inside the function a local variable?
```


#### **OTHER**

* I'm not sure if this was just on my side or whether it was intentional, but when I opened the .html file, a lot of the code were not in the chunks format. I also opened the .Rmd file to check it and the code was there, but in some parts it was not contained within the chunk format. I would have to manually put them in a chunk to be able to run them.
* Due to this, especially for the air quality data analysis part, since the code was not in the chunks, the outputs were not displayed in the .html file and I would need to run it in the .Rmd file to see the output.

  
  