---
title: "HW#4:Stringr & Air Quality Data"
author: "Vicky Espinoza"
date: "Due March 07, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
##R Book: Chapter 11 - Strings with stringr Exercises {.tabset}
<span style="color:red">1. Complete the tutorial steps for Strings from Chapter 14 in your R for Data Science text book http://r4ds.had.co.nz/strings.html. Answer all of the exercise questions in Chapter 14. Turn in your responses as R notebook file (.html) as a commit to your GitHub page. Turn in all of your functions as separate function (.R) files with intelligible titles as a commit to your GitHub page. </span>


##String Basics {.tabset}
<span style="color:blue">1. What is the difference between paste() and paste0()? What stringr function are they equivalent to? How do the functions differ in handling NA? </span>

paste () converts arguments to character strings, and concatenates them with separation specified by sep. By default it separates strings by spaces. It coerces missing value characters to NA. 

paste0() is more efficient than paste() at concantenating strings without spaces. NAs are taken as characters in the string. 

<span style="color:blue">2. Difference between sep and collapse arguments to str_c().</span>

sep is the element that separates every term in a string, whereas, collapse will return concatenated pairs as part of a single character vector with tuples separated by the string passed by collapse. 

<span style="color:blue"> 3.str_length() and str_sub() to extract the middle character from a string. what will you do if the string has an even number of characters? </span>


library(tidyverse)
library(stringr)
x <- c("abc", "abcd")
y <- str_length(x)
z <- ceiling(y / 2) 
str_sub(x, z, z)


When there are even number then the choice is arbituary and the ceiling or floor function is selected by the user in order to select the middle value. 

<span style="color:blue"> 4. What does str_wrap() do? When might you want to use it? </span>

This function wraps the text so that it fits within a certain width. Might want to use it when typing long strings of text.

<span style="color:blue"> 5.What does str_trim() do? What's the opposite of str_trim()? </span>

str_trim() trims the white space of a string and can specify the side that we want white space trimmed from, either left or right side as demonstrated below:

```{r}
library(tidyverse)
library(stringr)
str_trim(" Data! ")
str_trim(" Data! ", side = "left")
str_trim(" Data! ", side = "right")
```

The opposite of str_trim() is str_pad which adds characters to either side of the string. 
```{r}
library(tidyverse)
library(stringr)
str_pad("Data!",7, side = "both")
str_pad("Data!",6, side = "right")
str_pad("Data!",10, side = "left")
```

<span style="color:blue"> 6. Write a function that turns a vector c("a", "b", "c") into the string a, b, and c. Think carefully about what it should do if given a vector of length 0,1, or 2. </span>

```{r}
library(tidyverse)
library(stringr)
commasep <- function(x, sep = ", ", last = ", and ") {
  if (length(x) > 1) {
    str_c(str_c(x[-length(x)], collapse = sep),
                x[length(x)],
                sep = last)
  } else {
    x
  }
}
commasep("a")
commasep("data")
commasep(c("data", "stats", "analysis"))
```

##Matching Patterns with Regular Expressions {.tabset}
<span style="color:blue"> 1. Explain why each of these strings don't match a \: "\", "\\", "\\\".</span>

"\": escapes the next character in a R string
"\\": this creates a regular expression which will escape the next character in the string
"\\\": first two backlashes of this resolves the backlash of the regular expression and the third backlash escapes the next character in the string. it's the escape of a character in a regular expression

<span style="color:blue"> 2. How would you match the sequence "'\ ? </span>

a <- "\"'\\"
writeLines(a)
str_view(a, "\\\"'\\\\")


<span style="color:blue"> 3. What patterns will the regular expression \..\..\.. match? How would you represent it as a string? </span>

This expression will match strings with patterns that have a dot followed by a character (repeated 3x). 

To match it: 

a <- ".a.b.c"
writeLines(a)
str_view(a, "\\..\\..\\..")



##Anchors {.tabset}

<span style="color:blue"> How would you match the literal string `"$^$"`? </span>


string1 <- "$^$"
writeLines(string1)
str_view(string1, "\\$\\^\\$")



<span style="color:blue"> Given the corpus of common words in stringr::words, create regular expressions that find all words that: </span>

<span style="color:blue">a. Start with "y".</span>
str_view(words, "^y", match = TRUE)

<span style="color:blue">b. End with "X". </span>
str_view(words, "x$", match = TRUE)

<span style="color:blue">c.Are exactly three letters long.</span>
str_view(words, "^\\w{3}$", match = TRUE)
Note: This can also be done using str_length. 

<span style="color:blue">d. Have seven letters or more.</span>
str_view(words, "^\\w{7,}$", match = TRUE)

##Character Classes and Alternatives 

<span style="color:blue"> 1. Create regular expressions to find all words that:</span>

<span style="color:blue"> a.Start with a vowel </span>
str_view(words, "^[aeiouy]", match=T)

<span style="color:blue"> b. Contain only consonants (matching non-vowels) </span>
str_view(words, "^[^aeouiy]$", match = T)

<span style="color:blue">c. End with `ed`, but not with `eed`</span>
str_view(words, "[^e]ed$", match = T)

<span style="color:blue">d. End with `ing` or `ize`</span>
str_view(words, "ing$|ize$") 

<span style="color:blue"> 2. Empirically verify the rule "i before e except after c." </span>
str_view(words, "([^c]|)ei", match = T) # not always

<span style="color:blue">3.Is "q" always followed by a "u"?</span>
str_view(words, "q[^u]", match = T) 
YES!

4. Write a regular expression that matches a word if it;s probably written in British English, not American English.</span>
str_view(words, "ize", match = T)
american_british <- tibble::tribble(
  ~American.English, ~British.English,
            "color",         "colour",
         "behavior",      "behaviour",
          "theater",        "theatre",
         "behavior",      "behaviour",
            "meter",          "metre",
         "organize",       "organise",
         "traveled",      "travelled"
  )

str_view(american_british[[1]], "our$|re$|se$|lled")
str_view(american_british[[2]], "our$|re$|se$|lled")

<span style="color:blue"> Create a regular expression that will match telephone numbers as commonly written in your country.</span>

phone_numbers <- tibble::tribble(
                      ~phone_numbers,
                    "(818) 257-1201",
                    "(818)-257-1201",
                      "818-257-1201",
                          "257-1201",
                          "257 1201", 
                          "2571201")
str_view(phone_numbers[[1]],"^\\(?\\d{3}\\)?[\\s|-](\\d{3}[\\s|-]{1}\\d{4}$|\\d{4}$)")

##Repetition {.tabset}
<span style="color:blue">1. Describe the equivalents of ?, +, * in {m,n} form.</span>
?: {0,1}
+ : {1,}
* : {0,}


<span style="color:blue">2.Describe in words what these regular expressions match (read carefully to see if I'm using a regular expression or a string that defines a regular expression): </span>

<span style="color:blue">a.^.*$</span>
Matches any character from beginning of the string to the end of string. Matches any input. 

<span style="color:blue">b."\\{.+\\}"</span>
Matches a backlash, open curly bracket, any character, backlash and closed curly bracket.

<span style="color:blue">c.\d{4}-\d{2}-\d{2}</span>
Matches 4 digits, a dash, 2 digits, a dash, and 2 digits 

<span style="color:blue">d."\\\\{4}"</span>
Matches \\\\

<span style="color:blue">3. Create regular expressions to find all the words that: </span>

<span style="color:blue">a. Start with three consonants.</span>
str_view(words, "^[^aeouiy]{3}", match = T)

<span style="color:blue">b.Have 3+ vowels in a row </span>
str_view(words, "[aeouiy]{3,}", match = T)

<span style="color:blue">c.Have 2+ vowel-constant pairs in a row </span>
str_view(words, "([aeouiy][^aeouiy]){2,}", match = T)


<span style="color:blue">4. Solve the beginner regexp crosswords at https://regexcrossword.com/challenges/beginner </span>

<span style="color:blue">1. Beatles </span>
TOP:H,E 
BOTTOM:L,P
<span style="color:red">HELP </span>

<span style="color:blue">2. Naughty </span>
TOP:B,O
BOTTOM:B,E
<span style="color:red"> BOBE </span>

<span style="color:blue">3. Ghost </span>
TOP:O,O
BOTTOM:O,O
<span style="color:red"> OOOO </span>

<span style="color:blue">4.Symbolism </span>
TOP:*,*
BOTTOM:/,/
<span style="color:red"> **//  </span>

<span style="color:blue">5. Airstrip One </span>
TOP:1,9
BOTTOM:8,4
<span style="color:red"> 1894 </span>

##Grouping and Backreferences {.tabset}
<span style="color:blue">1. Describe, in words, what these expressions will match: </span>
<span style="color:blue">1. (.)\1\1</span>
Results in 3 characters 3 times in a row

<span style="color:blue">2. "(.)(.)\\2\\1" <span>
A pair of two characters followed by the same pair of characters in reverse order 

<span style="color:blue">3.(..)\1 </span>
Two characters repeated

<span style="color:blue">4. "(.).\\1.\\1" <span>
 A character (original) followed by any character, the original character, any other character, the original character

<span style="color:blue">5. "(.)(.)(.).*\\3\\2\\1" <span>
This matches 3 characters followed by 0+ characters (any kind) followed by the same three characters in reverse order. 

<span style="color:blue">2.Construct regular expressions to match words that: </span>

<span style="color:blue">Construct regular expressions to match words that: </span>

<span style="color:blue">a.Start and end with the same character.</span>
^(.).*\1$
str_view(words, "^(.).*\\1$")

<span style="color:blue">b.Contain a repeated pair of letters </span>
"(..).*\\1"
str_view(words, "(..).*\\1")

<span style="color:blue">c.Contain one letter repeated in at least three places </span>
"([a-z]).*\\1.*\\1"

##Tools: Detect Matches {.tabset}
<span style="color:blue">1.For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls: </span>

<span style="color:blue">a.Find all words that start or end with x.</span>

```{r}
start_with_x <- str_detect(words, "^x")
end_with_x <- str_detect(words, "x$")
words[start_with_x | end_with_x]
```

<span style="color:blue">b.Find all words that start with a vowel and end with a consonant.</span>
```{r}
str_subset(words, "^[aeiou].*[^aeiou]$") %>% head()
start_with_vowel <- str_detect(words, "^[aeiou]")
end_with_consonant <- str_detect(words, "[^aeiou]$")
words[start_with_vowel & end_with_consonant] %>% head()
```

<span style="color:blue">c. Are there any words that contain at least one of each different vowel? </span> No there are no words that contain at least one of each different vowel. 
```{r}
library(stringr)
words[str_detect(words, "a") &
        str_detect(words, "e") &
        str_detect(words, "i") &
        str_detect(words, "o") &
        str_detect(words, "u")]
```

<span style="color:blue">d.What word has the highest number of vowels? What word has the highest proportion of vowels?(Hint what is the denominator) </span>

```{r}
vowels<- str_count(words, "[aeiou]") / str_length(words)
words[which(vowels == max(vowels))]
```
Out of the words dataset in stringr it states that the word with the highest number of vowels is a. 

 
##Tools: Extract Matches {.tabset}
<span style="color:blue">1. In the previous example, you might have noticed that the regular expression matched flickered, which is not a color. Modify the regex to fix the problem. </span>

<span style="color:green">This was the previous expression:
```{r}
colors <- c("red", "orange", "yellow", "green", "blue", "purple")
color_match <- str_c(colors, collapse = "|")
sentence<- sentences[str_count(sentences,color_match)>1]
sentence
```

The word flickered because it has the color red embedded in the word. 

```{r}
colors <- c("red", "orange", "yellow", "green", "blue", "purple")
color_match<- str_c("\\b(", str_c(colors, collapse = "|"), ")\\b")
sent<- sentences[str_count(sentences,color_match)>1]
sent
```
<span style="color:blue">2.From the Harvard sentence data, extract: </span>
<span style="color:blue">a.The first word from each sentence</span>

```{r}
str_extract(sentences, "[a-zA-X]+") %>% head()
```

<span style="color:blue">b. All words ending in -ing </span>
```{r}
pattern <- "\\b[A-Za-z]+ing\\b"
ing_sentence <- str_detect(sentences, pattern)
unique(unlist(str_extract_all(sentences[ing_sentence], pattern))) %>%
  head()
```
<span style="color:blue">c. All plurals </span>
```{r}
unique(unlist(str_extract_all(sentences, "\\b[A-Za-z]{3,}s\\b"))) %>%
  head()
```

##Tools: Grouped Matches {.tabset}
<span style="color:blue">1.Find all words that come after a "number" line "one" "two", "three", etc. Pull out both the number and the word. </span>
```{r}
num<- "(one|two|three|four|five|six|seven|eight|nine|ten) +(\\S+)"
sentences[str_detect(sentences, num)] %>%
  str_extract(num)
```

<span style="color:blue">2.Find all contractions. Separate out the pieces before and after the apostrophe. </span>
```{r}
contraction <- "([A-Za-z]+)'([A-Za-z]+)"
sentences %>%
  `[`(str_detect(sentences, contraction)) %>%
  str_extract(contraction)
```

##Tools: Replacing Matches {.tabset}
<span style="color:blue">1. Replace all forward slashes in a string with backslashes </span>
```{r}
backslash <- str_replace_all("Desktop/vespinoza/Desktop/ENVE207", "\\/", "\\\\")
writeLines(backslash)
```

<span style="color:blue">2. Implement a simple version of str_to_lower() using replace_all() </span>
```{r}
word<- c("HELLO")
lowercase <-str_replace_all(word, c("A"="a", "B"="b", "C"="c", "D"="d", "E"="e", "F"="f", "G"="g", "H"="h", "I"="i", "J"="j", "K"="k", "L"="l", "M"="m", "N"="n", "O"="o", "P"="p", "Q"="q", "R"="r", "S"="s", "T"="t", "U"="u", "V"="v", "W"="w", "X"="x", "Y"="y", "Z"="z"))
lowercase
```
<span style="color:blue">3. Switch the first and last letters in words. Whcih of those strings are still words? </span>
```{r}
switch <- str_replace_all(words, "^([A-Za-z])(.*)([a-z])$", "\\3\\2\\1")
still_words <-intersect(switch,words)
still_words
```
45 words still remain words after swapping the first and last letters of words. 

##Tools: Splitting
<span style="color:blue">1.Split up a string like "apples, pears, and bananas" into individual components.</span>
```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
```

<span style="color:blue">2.Why is it better to split up by boundary("word") than " "? </span>
Splitting on boundary("word") splits on both white space and punctuation.

<span style="color:blue">3.What does splitting with an empty string ("") do? Experiment, and then read the documentation. </span>
```{r}
str_split("Vi. ck|y", "")[[1]]
```
It splits the string into individual characters. 

##Other Types of Patterns {.tabset}
<span style="color:blue">1.How would you find all strings containing \ with regex() versus with fixed () ? </span>
```{r}
str_subset(c("da\\ta", "data", "data\\sets"), "\\\\")
str_subset(c("da\\ta", "data", "data\\sets"), fixed("\\"))
```

<span style="color:blue">What are the five most common words in sentences? </span>

```{r}
library(stringr)
str_extract_all(sentences, boundary("word")) %>%
  unlist() %>%
  str_to_lower() %>%
  tibble() %>%
  set_names("word") %>%
  group_by(word) %>%
  count(sort = TRUE) %>%
  head(5)
```
The five most common words in sentences are: 1. the, 2. a, 3. of, 4. to, 5. and

##stringi {.tabset}
<span style="color:blue">1. Find the stringi functions that: </span>

<span style="color:blue">a. Count the number of words </span>
stri_count_words 

<span style="color:blue">b. Find duplicated strings </span>
stri_duplicated 

<span style="color:blue">c.Generate random text </span>
stri_rand_
stri_rand_strings (generates random strings)

##Air Quality Data Analysis {.tabset}
<span style="color:blue">2. What class is o3.filelist? What does it contain?> </span>

o3.filelist is class "list" and contains tibbles (dataframes) compiled from the .txt files 


<span style="color:blue">3. Using ~ 1 sentence per line in the above code, explain what each line in the code is doing. </span>

o3.filenames <- list.files(pattern = ".txt") <span style="color:green"> In element o3.filenames store a vector of the names of files located in the specified directory using ".txt" pattern to find the names of the files </span> 

o3.filelist <- lapply(o3.filenames, read_delim, delim = "|") <span style="color:green"> In element o3.filelist store the information delimited by a "|" that is located in each of the text files with the filenames located in o3.filenames as tibbles. </span>

names(o3.filelist) <- gsub(".txt","", o3.filenames)  <span style="color:green"> Replacing the name of the o3.filelist  ".txt" portion of the o3.filenames with " ". Where o3.filenames are what are being sought. </span>

<span style="color:blue"> 4. Rewrite the code above using the stringr package instead of grep{base}.</span>

names(o3.filelist) <- str_replace_all(o3.filenames,".txt", "")

<span style="color:blue"> 5.Rewrite the code above using traditional object-based R programming instead of the piping operator.</span>


binding <- rbindlist(o3.filelist)
group <- group_by(binding,site = as.factor(site), date)
sum_o3_other <- summarize(group,o3 = mean(obs, na.rm = TRUE))
sum_o3_other
identical(sum_o3_other,daily)


<span style="color:blue"> 6. Summarize the o3 data above by site and by month and by year using a piping operator (the monthly mean o3 for each site for each year).</span>


monthly <- o3.filelist %>%
  rbindlist() %>%
  group_by(site = as.factor(site), year(date), month(date)) %>%
  summarize(o3 = mean(obs, na.rm = TRUE))
monthly


<span style="color:blue">7. Challenge! Ozone pollution actually follows a very strong diurnal pattern. How would you summarize the daily data from above in a better way to capture that diurnal pattern?</span>

hourly<- o3.filelist %>%
  rbindlist() %>%
  group_by(site = as.factor(site), start_hour, year(date)) %>%
  summarize(o3 = mean(obs, na.rm = TRUE))
#head(hourly)
ggplot(data= hourly)+
  geom_point(mapping=aes(x=start_hour,y=o3))


<span style="color:blue">8. How many site names in the CA air quality location dataset "Site Name" contain "San" or "Santa?".</span>

#this one does not take into account embedded duplicates whereas the second code addresses this via the collapse
san_sum <- sum(str_count(loc$'Site Name',"San"))
santa_sum <- sum(str_count(loc$'Site Name',"Santa"))
total <- sum(santa_sum,san_sum)
total

#or 
x <- c("San", "Santa")
name_match <- str_c(x,collapse = "|")
name_subset <- str_subset(loc$`Site Name`,name_match)
count_names <- str_count(name_subset,name_match)
sum(count_names)


<span style="color:blue">9. Identify the number of sites that do not have a complete address (full street address and zip code).</span> 

The total number of sites that do not have a complete address and zip code are 262

sum(duplicated(loc$Location)) #this is to confirm that Locations are not duplicated 
df <- tibble(
  address=loc$Address, Zip=loc$`Zip Code`
)
sum(is.na(df))


<span style="color:blue">10. How does a semi join differ from an inner join?</span>
A semi join is different from an inner join in that the inner join will return one row of x for each matching row of y, whereas the semi join will never duplicate rows of x.

<span style="color:blue">11. Write a function to calculate the annual mean, median, max and min of all sites that have "San" or "Santa" in their name.</span>

#Need to extract all the rows that have data for San and Santa 

I was able to get a result for this, but rerunning the code was not functioning. I will have to go back and debug. Here is my attempt. 

library(data.table)
library(tidyverse)

site.summary <- function(x,y,z){
  yearly <- x %>%
    group_by(site = as.factor(site))%>%
    mutate(year = format(date, "%Y")) %>%
    group_by(site,year) %>%
    c <- c(y, z)
  name_match <- str_c(c,collapse = "|")
  d <- str_subset(loc$`Site Name`,name_match)
  San <- subset(yearly, yearly$`Site Name`== d )%>%
    group_by(site,year )%>%
    summarize(mean=mean(o3, na.rm=TRUE),median=median(o3, na.rm = TRUE),max=max(o3, na.rm=TRUE),min=min(o3, na.rm=TRUE))
  San
}
site.mean(daily.site,"San","Santa")



<span style="color:blue">12. Write a function to caculate the annual daily mean. Apply that function to Merced County. What is the annual daily mean of o3 for Merced County?</span>
I was able to get a result for this, but rerunning the code was not functioning. I will have to go back and debug. Here is my attempt. 

library(data.table)
library(tidyverse)
site.mean <- function(x){
  x <- c("Merced")
  name_match <- str_c(x,collapse = "|")
  name_subset <- str_subset(loc$`Site Name`,name_match)
  merceddata <- subset(daily.site, daily.site$`Site Name`== name_subset)
  mean=mean(x, na.rm=TRUE)
}
site.mean(merceddata$o3)
mean










  

